---
title: "Making the stable cause weaker"
output:  
  html_document:
    number_sections: true
    toc: true  
    collapsed: false
    toc_float: true
    smooth_scroll: false
    toc_depth: 3
---

```{r setup, include=FALSE}
# packages
library(ez)
library(reshape2)
library(reshape)
library(ggplot2)
library(plyr)
library(pastecs)
library(ez)
library(data.table)
library(tidyverse) 

library(showtext)
library(readr)


font_add_google("Poppins", "Poppins")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()
```


```{r}
tdata <- read_delim("Exp3_data.txt", "\t", escape_double = FALSE, trim_ws = TRUE)

N <- table(tdata$causal_strength, tdata$decision_level)
N

N_gov <- sum(N[,1])
N_pers <- sum(N[,2])

table(tdata$condition_1)

```


# Subject demographics

```{r}
# demographics 

min(tdata$age)
max(tdata$age)
mean(tdata$age)
sd(tdata$age)

# 1 = male, 2 = female, 3 = other
table(tdata$gender)
```

1 = male, 2 = female, 3 = non-binary, 4 = prefer not to say


# Results

```{r, echo = FALSE}
# reorder factor 

# to create chart, the data must be in long format and only contain the relevant dependent variables

# 1. make a subset with only the relevant dvs 

tdata_sub <- subset(tdata, select = c(run_id,causal_strength, decision_level, main_dv_rec))


tdata_long <- tdata_sub
tdata_long$causal_strength <- factor(tdata_long$causal_strength, levels = c("eightyfive", "seventyfive", "sixtyfive"), 
                               labels = c("mean causal strength: \n0.90 vs. 0.85", "mean causal strength: \n0.90 vs. 0.75",
                                          "mean causal strength: \n0.90 vs. 0.65"))


tdata_long$decision_level <- factor(tdata_long$decision_level, levels = c("general_norm", "personal"), 
                              labels = c("general norm", "personal"))


```

## Graphs

```{r}
myTheme <- theme(plot.title = element_text(face="bold", size = 22),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 16, angle = 0), 
        axis.text.y = element_text(size = 16, angle = 0),
        legend.text = element_text(size = 18),
        legend.title = element_text(face = "bold", size = 18),
        strip.text.x = element_text(size = 18),
        #panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line.x = element_line(colour = "black"), 
        axis.line.y = element_line(colour = "black"),
        axis.text = element_text(colour ="black"), 
        axis.ticks = element_line(colour ="black"))

tdata_sub <- tdata_long


library(see)
## first, turn sID into a factor
tdata_sub$run_id <- factor(tdata_sub$run_id)

pd <- position_dodge(width = 0.3)

tdata_sub$valueJitter <- jitter(tdata_sub$main_dv_rec, factor = 1, amount = 0.04)

theme_set(theme_light(base_size = 20, base_family = "Poppins"))

# new labes for the facets 

g <- ggplot(tdata_sub, aes(x=decision_level, y=valueJitter)) +
  guides(fill=FALSE)+
  facet_grid( ~ causal_strength)+
  #ggtitle("Subjects' causal srength ratings") +
  scale_y_continuous(limits = c(-0.05, 1.05), breaks=seq(0, 1, 0.1), expand = c(0,0)) +
  #scale_x_discrete(labels=c("Single-effect \ncause", "Common \ncause", "No \ncause")) +
  #stat_summary(fun.y = mean, geom = "bar", position = "dodge", colour = "black", alpha =0.5) +
  geom_violinhalf(aes(y = main_dv_rec, group = decision_level, fill = decision_level), color = NA, position=position_dodge(1), alpha = 0.4)+
  #geom_line(position = pd, color = "black", size = 1, alpha=0.04) +
  geom_hline(yintercept=0.5, linetype="dashed", color = "black")+
  geom_jitter(aes(color = decision_level), alpha = 0.5, width = 0.1) +
  stat_summary(aes(y = main_dv_rec,group=1), fun.data = mean_cl_boot, geom = "errorbar", width = 0, size = 1) +
  stat_summary(aes(y = main_dv_rec,group=1, fill = decision_level), fun.y=mean, geom="point", color = "black", shape = 22, size = 5, group=1, alpha = 1)+
  stat_summary(aes(y = main_dv_rec,group=1), fun.y=median, geom="point", color = "black", shape = 3, size = 4, group=1, alpha = 1, position = position_dodge(width = 0.5))+
  labs(x = "Context (decison Level)", y = "Causal preference") +
  scale_color_manual(name = "Dec. Level",values=c("#66c2a5", "#e78ac3"))+
  scale_fill_manual(name = "Dec. Level",values=c("#66c2a5", "#e78ac3"))+
  theme(legend.position = "none")+
  myTheme
g


#ggsave("results_means.svg",width=11,height=5)
#ggsave("results_means.pdf",width=11,height=5)
```





## Analyses

```{r, echo = FALSE, warning = FALSE, message = FALSE}

################################################################################################################
##################################### Statistical Analyses #####################################################
################################################################################################################
library(pastecs)
library(lme4)
library(nlme)
library(ez)
library(vtable)
```
Descriptive stats: 

```{r}
sum_data <- tdata_sub
sum_data$cond <- paste(sum_data$causal_strength, sum_data$decision_level)

sum <- st(sum_data, vars = c('main_dv_rec'), group = c("cond"), add.median=T, out='return', group.long=T)
sum

# see: https://cran.r-project.org/web/packages/vtable/vignettes/sumtable.html
```

### ANOVA 


```{r}
library(afex)
library(emmeans)

a1 <- aov_car(main_dv_rec ~ decision_level*causal_strength + Error(run_id), tdata_sub, anova_table = list(es = "pes"))
a1


```


### Contrasts

```{r}
############### 
# a follow-up analysis 

# means

ls2 <- lsmeans(a1, c("decision_level", "causal_strength")) # group means by between-condition
ls2

# contrast the decision levels (main effect)
contrasts <- emmeans(a1, ~ decision_level)
d <- pairs(contrasts, adjust = "none")


# contrast the strength levels (main effect; averaging over decision level, as there was no sig. interaction)
contrasts <- emmeans(a1, ~ causal_strength)
s <- pairs(contrasts, adjust = "none")

d
confint(d, level = 0.95)

s
confint(s, level = 0.95)
```
Polynomial contrast to test for global trends (mean ratings should decrease constantly)

```{r}
# all contrasts were significant, but a better way than doing three tests is to do a polynomial trend test

# Polynomial Contrasts
#library(Rcmdr)
# Quantitative Trend analysis

contrasts(tdata_sub$causal_strength) <- "contr.poly" # define contrast coefficients

Trend <- aov(main_dv_rec ~ causal_strength, data=tdata_sub)
summary.lm(Trend)



```

Only a significant negative linear trend. 


### Standardized effect sizes

What Cohen's d is the main effect for decision level?


```{r}

# strength = 0.90

dat <- tdata_sub



# now compute ES and SE and CI of it
# using the esc package because it gives SE of the ES directly
library(esc)

# get means and sds
m1 <- dat %>%
          filter(decision_level == "general norm")%>%
          summarize(Mean1 = mean(main_dv_rec))

sd1 <- dat %>%
          filter(decision_level == "general norm")%>%
          summarize(SD1 = sd(main_dv_rec))

n1 <-  nrow(subset(dat, decision_level == "general norm"))


m2 <- dat %>%
          filter(decision_level == "personal")%>%
          summarize(Mean2 = mean(main_dv_rec))

sd2 <- dat %>%
          filter(decision_level == "personal")%>%
          summarize(SD2 = sd(main_dv_rec))

n2 <-  nrow(subset(dat, decision_level == "personal"))




d <- esc_mean_sd(
  grp1m = m1[,1], grp1sd = sd1[,1], grp1n = n1,
  grp2m = m2[,1], grp2sd = sd2[,1], grp2n = n2,
  es.type = "d"
)
d

d$ci.lo
d$ci.hi

d_ci <- (d$ci.hi - d$ci.lo)/2
d_ci
```





## Cluster Analysis 

### Clustering

```{r}

crit = 0
tdata_long$Cluster <- 0
tdata_long$Cluster[tdata_long$main_dv_rec > 0.50] <- "Stable Cause"
tdata_long$Cluster[tdata_long$main_dv_rec < 0.50] <- "Unstable Cause"
tdata_long$Cluster[tdata_long$main_dv_rec == 0.50] <- "Indifferent"


```


```{r}
# create a summary dataset that also contains the percentages
tdata_sub <- tdata_long
plotdata_between <- tdata_sub %>%
  group_by(causal_strength, decision_level, Cluster) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))
```


### Proportion tests 

Conducted to get the CIs of the proportion estiamtes

```{r}
# append CI of the proportions 
cases <- plotdata_between$n


t <- prop.test(cases[1],N_gov/3,correct=FALSE) # get CI for Invariance government
CI_gov_St9085_min <- t$conf.int[1]
CI_gov_St9085_max <- t$conf.int[2]


t <- prop.test(cases[2],N_gov/3,correct=FALSE) # get CI for Invariance government
CI_gov_USt9085_min <- t$conf.int[1]
CI_gov_USt9085_max <- t$conf.int[2]


t <- prop.test(cases[3],N_pers/3,correct=FALSE) # get CI for Invariance government
CI_pers_St9085_min <- t$conf.int[1]
CI_pers_St9085_max <- t$conf.int[2]

t <- prop.test(cases[4],N_pers/3,correct=FALSE) # get CI for Invariance government
CI_pers_USt9085_min <- t$conf.int[1]
CI_pers_USt9085_max <- t$conf.int[2]



t <- prop.test(cases[5],N_gov/3,correct=FALSE) # get CI for Invariance government
CI_gov_St9075_min <- t$conf.int[1]
CI_gov_St9075_max <- t$conf.int[2]


t <- prop.test(cases[6],N_gov/3,correct=FALSE) # get CI for Invariance government
CI_gov_USt9075_min <- t$conf.int[1]
CI_gov_USt9075_max <- t$conf.int[2]


t <- prop.test(cases[7],N_pers/3,correct=FALSE) # get CI for Invariance government
CI_pers_St9075_min <- t$conf.int[1]
CI_pers_St9075_max <- t$conf.int[2]

t <- prop.test(cases[8],N_pers/3,correct=FALSE) # get CI for Invariance government
CI_pers_USt9075_min <- t$conf.int[1]
CI_pers_USt9075_max <- t$conf.int[2]



t <- prop.test(cases[9],N_gov/3,correct=FALSE) # get CI for Invariance government
CI_gov_St9065_min <- t$conf.int[1]
CI_gov_St9065_max <- t$conf.int[2]


t <- prop.test(cases[10],N_gov/3,correct=FALSE) # get CI for Invariance government
CI_gov_USt9065_min <- t$conf.int[1]
CI_gov_USt9065_max <- t$conf.int[2]


t <- prop.test(cases[11],N_pers/3,correct=FALSE) # get CI for Invariance government
CI_pers_St9065_min <- t$conf.int[1]
CI_pers_St9065_max <- t$conf.int[2]

t <- prop.test(cases[12],N_pers/3,correct=FALSE) # get CI for Invariance government
CI_pers_USt9065_min <- t$conf.int[1]
CI_pers_USt9065_max <- t$conf.int[2]




CI_low <- c(CI_gov_St9085_min, CI_gov_USt9085_min, 
            CI_pers_St9085_min, CI_pers_USt9085_min,
            CI_gov_St9075_min, CI_gov_USt9075_min, 
            CI_pers_St9075_min, CI_pers_USt9075_min,
            CI_gov_St9065_min, CI_gov_USt9065_min, 
            CI_pers_St9065_min, CI_pers_USt9065_min)


CI_up <- c(CI_gov_St9085_max, CI_gov_USt9085_max, 
            CI_pers_St9085_max, CI_pers_USt9085_max,
            CI_gov_St9075_max, CI_gov_USt9075_max, 
            CI_pers_St9075_max, CI_pers_USt9075_max,
            CI_gov_St9065_max, CI_gov_USt9065_max, 
            CI_pers_St9065_max, CI_pers_USt9065_max)

plotdata_between$CI_low <- CI_low
plotdata_between$CI_up <- CI_up

```


### Graph

```{r}
library(scales)
theme_set(theme_light(base_size = 10, base_family = "Poppins"))

plotdata <- plotdata_between
#plotdata <- subset(plotdata, Change != "Invariance")

g<- ggplot(plotdata, 
       aes(x = decision_level,
           y = pct,
           fill = Cluster)) +
  facet_grid( ~ causal_strength)+
  geom_bar(stat = "identity",
           position = "dodge") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  coord_cartesian(xlim =c(1, 2), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  #geom_text(aes(label = lbl), 
  #          size = 4, 
  #          position = position_stack(vjust = 0.5)) +
  geom_text(aes(label = lbl), 
            size = 4, 
            position = position_dodge(width = .9),
            vjust = +2.2) +
  geom_errorbar(aes(ymin=CI_low, ymax=CI_up), width=.0,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Pastel1") +
  scale_fill_manual(values=alpha(c("#fc9272", "#3182bd"),0.7)) +
  labs(y = "Percentage", 
       fill = "Preference",
       x = "Context (Decision Level)")+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  #annotate("pointrange", x = c(1,2), y = 1 - plotdata$pct, 
  #         ymin = 1 - plotdata$pct - plotdata$CI, 
  #         ymax = 1- plotdata$pct + plotdata$CI, 
  #         colour = "black", size = 0.8, shape = 22, fill = "lightblue", fatten = 1)+
  #annotate("text", x = 1.25, y = Inf, label = "p < .001", size = 4, vjust = 1.6)+
  #annotate("text", x = 2.35, y = Inf, label = "p = .09", size = 4, vjust = 4)+
  theme(legend.position = "top", axis.title = element_text(size = 12), axis.text = element_text(size = 10, color = "black"),
        legend.text = element_text(size = 10),legend.title = element_text(size = 12),
        strip.text.x = element_text(size =11))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g
#ggsave("selections_Dim.pdf",width=7,height=4)
#ggsave("selections_Dim.svg",width=7,height=4)
```




